Formal Analysis of EDHOC Key Establishment for Constrained IoT Devices

### 2022-07-05 paper.tex
[add] update of DoS section

>>>>

%-------------------------------------------------------------------------- sec
\section{\uppercase{Denial of Service}}
\label{sec:dos}
%\mPoint{DoS is out of scope for \mEdhoc{}; apps to deal with it}
\mEdhoc{} includes no specific measures for countering Denial of Service (DoS)
attacks and delegates the responsibility to handle this to the application.
%
\mEdhoc{} itself has limited functionality, few negotiable parameters and
processing rules.
%
Hence, we argue in this section that its attack surface
is limited from a DoS perspective.
%
However, \mEdhoc{} leaves much of the semantics of its messages undefined.
%
This increases the risk of applications vulnerable to DoS via \mEdhoc{}, even
though \mEdhoc{} itself is arguably not particularly weak in this respect.
%

\paragraph{DoS mitigation is left to developers.}
Section 8.7 of the \mEdhoc{} specification\footnote{draft-ietf-lake-edhoc-14}
explains the design philosophy w.r.t.
DoS in two paragraphs and gives two recommendations.
%
The first being that \mEdhoc{} relies on lower layers to mitigate DoS, and an
example of DoS countermeasures is given: checking return routeability when
receiving the first message.
%
The second being that an application may try to determine (in some way) that an
apparently valid message is in fact probably a forgery and should be ignored or
cause processing to stop.
%

No recommendation is given on how to minimize DoS effects stemming from use of
error messages, which can be used in an arbitrary way in \mEdhoc{}.
%
We now elaborate on some considerations application developers needs to take
into account.
%

\paragraph{Amplification and asymmetry.}
For the same reasons as given in the IKEv2 DDoS analysis
by~\cite{rfc8019} the responder is more vulnerable than the
initiator.
%
The key similarity between IKEv2 and \mEdhoc{} here is that the initiator is
anonymous in the first message and that creating the first message is
computationally cheap and requires no storage.
%
No element in the first \mEdhoc{} message can be verified by the receiver as
being legitimate as long as the format is correct and elliptic curve points
lie on the proposed curve.
%
\mEdhoc{} leave it to the application and its
transport layer to resolve any issues.
%
In addition to the computational resources the responder spends, communication
patterns in the underlying protocols may require the responder to
obtain and verify certificate chains, and verify
certificate-revocation servers as a result of the spoofed initial message.
%
For example, the content of the \mADone{} information element may hint the
responder to take further actions.
%

The asymmetry between the adversary's effort and the responder's may hence be
significant.
%
Since \mEdhoc{} specifically targets constrained devices this asymmetry needs to
be considered even more carefully than for IKEv2~\cite{rfc8019}.
%

Depending on the use case, the first \mEdhoc{} message may be sent by either
party in a session.
%
For example, even though it may seem intuitive that an electronic
key-card initiates the connection towards lock, the first \mEdhoc{} message
could be included in the response from the lock.
%
This means the lock can postpone heavy verification actions until it has
authenticated the key card after receiving the third \mEdhoc{} message.
%

Implementers are well advised to consult \cite{rfc8019}, which provides much
good general information about DoS for protocols with this structure.
%
However, ultimately, a product-specific risk assessment is required to
determine appropriate DDoS mitigations.
%

\paragraph{Reflection Attacks}
Reflection attacks are attacks where an adversary spoofs the source address of a
victim in a message and thereby tricks the responder to send a message to the
victim.
%
\mEdhoc{} leaves protection against this to the transport protocol and
application, and suggests using a return path reachability method similar to IKEv2.
%
We note that such reachability mechanisms can easily be used to make a responder
send reachability requests to arbitrary addressable targets.
%
Depending on the design of such reachability mechanisms, they may require that
unsolicited reachability requests are discarded, limiting the effects of
anonymous DoS attacks like these.
%

In situations where the responder does not need to perform additional
communications to verify certificates or similar, a reachability mechanism may
be more expensive in terms of time and storage compared to continue with the
protocol and keeping a half-open state for a period of time.
%
Doing so would also avoid creating secondary DoS effects on certificate storage
nodes or certificate revocation servers.
%
The latter may cause a systems-level issue if many distributed
\mEdhoc{} responders frequently requests such servers.
%
However, if round-trip times are large, and half-open states hence are kept
by the receiver for a long time, a reachability mechanism can trade
communication for storage.
\footnote{RFC8019 discuss some numbers as examples of tradeoffs. Should we set
up a simple model and derive an expression for trade-offs? Could be useful for
implementors to consider, but low academic value. This would capture also
amplification.}
%

\mPoint{In a constrained device randomness may be easily depleted}
Assume an adversary generate (deterministic) bogus message 1s and send these at
a sufficiently high rate to the responder.
%
\mEdhoc{} specifically targets constrained devices, that presumably may not
always have good randomness sources.
%
Is it possible to deplete or significantly reduce the entropy of the source and
in this way be able to predict the responder's random input to other sessions
(with other initiators)?
%
The LAKE WG have been talking about effects of poor randomness on signatures, so
this should be an issue, no?
%

\mPoint{\mEdhoc{} can deadlock if not used correctly}
An adversary may send forged error messages, and specifically reject proposed
ciphersuites.
%
The initiator is recommended to not try that ciphersuite with the responder
again in a new session for a while.
%
If the adversary rejects all options for a initiator this way, it can prevent
communication between the initiator and that responder until the cache is
cleared.
%
Potentially, an adversary could lock a party of a group out of communication all
together this way.
%
This needs to be considered by applications.
%

While we have not conducted a formal analysis of the fact, the structure of all
methods are so similar that it seems plausible that none of them would be
stronger against DDoS than the others.
%
